---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

```{r, loadLibraries}
library(stringr)
```

Download the HDF5 source tarball and unpack into a temporary directory.  We rename the output folder to *hdf5*.

```{r}
hdf5_source <- tempfile()
download.file(url = "https://s3.amazonaws.com/hdf-wordpress-1/wp-content/uploads/manual/HDF5/HDF5_1_10_5/hdf5-1.10.5.tar.bz2", 
              dest = hdf5_source)
untar(tarfile = hdf5_source, exdir = tempdir())
system2("mv", args = c(file.path(tempdir(), "hdf5-1.10.5"), file.path(tempdir(), "hdf5")))
```


```{r}
hdf5_dir <- file.path(tempdir(), "hdf5")
```

Delete folders we are not distributing.

```{r}
unlink(x = file.path(hdf5_dir, c("examples", "fortran", "java", 
             "release_docs", "test", "testpar", "tools", 
             "c++/examples", "c++/test", "hl")),
       recursive = TRUE)
```

Next we make some modifications to configure scripts and make files, removing references to the folders we have deleted.

```{r}
configure_ac <- xfun::read_utf8(file.path(hdf5_dir, "configure.ac"))

## modify list of build files
start <- which(str_detect(configure_ac, pattern = "AC_CONFIG_FILES"))
end <- which(str_detect(configure_ac[start:(length(configure_ac))], pattern = "\\)$"))[1] + start - 1
config_files <- configure_ac[start:end]
rm_idx <- which(str_detect(config_files, pattern = "test/|testpar/|tools/|examples/|fortran/|java/|h5c++|hl/"))
config_files <- config_files[-rm_idx]
config_files[length(config_files)] <- paste0(tail(config_files, 1), "])")
configure_ac[start] <- paste(config_files, collapse = "\n")
configure_ac <- configure_ac[-((start+1):(end))]

## remove reference to h5cc
h5cc <- str_which(configure_ac, pattern = "chmod 755 tools/src/misc/h5cc")
configure_ac <- configure_ac[-((h5cc):(h5cc+4))]

## fortran headers
fortran_inc <- str_which(configure_ac, pattern = "AC_CONFIG_HEADERS\\(\\[fortran/src/H5config_f\\.inc")
configure_ac[fortran_inc:(fortran_inc+1)] <- paste("##", configure_ac[fortran_inc:(fortran_inc+1)])

## write 
xfun::write_utf8(configure_ac, con = file.path(hdf5_dir, "configure.ac"))

## C++ makefile
make_cplusplus <- xfun::read_utf8(file.path(hdf5_dir, 'c++/Makefile.am'))
idx <- str_which(make_cplusplus, "BUILD_CXX_CONDITIONAL")
make_cplusplus[idx] <- "if BUILD_CXX_CONDITIONAL\n   SUBDIRS=src\nendif\nDIST_SUBDIRS = src"
make_cplusplus <- make_cplusplus[-((idx+1):(length(make_cplusplus)-2))] 
xfun::write_utf8(make_cplusplus, con = file.path(hdf5_dir, "c++/Makefile.am"))

## Primary makefile
make <- xfun::read_utf8(file.path(hdf5_dir, 'Makefile.am'))
idx <- str_which(make, "SUBDIRS = src test")[1]
make[idx] <- "SUBDIRS = src . $(CXX_DIR)"
make[idx+1] <- "DIST_SUBDIRS = src . c++"
make[idx+2]  <- ""
idx <- str_which(make, "# Make all, tests, and \\(un\\)install")
make[(idx+1):(idx+6)] <- paste0("##", make[(idx+1):(idx+6)])
xfun::write_utf8(make, con = file.path(hdf5_dir, "Makefile.am"))
```

Address problem in source code reported on CRAN Solaris builder.

```{r}
code <- xfun::read_utf8(file.path(hdf5_dir, 'c++', 'src', 'H5Library.cpp'))
code <- str_replace(code, '([ ]{1,})(exit\\()', replacement = '\\1std::\\2' )
xfun::write_utf8(code, con = file.path(hdf5_dir, 'c++', 'src', 'H5Library.cpp'))
```

Process the modified scripts.

```{r}
system(command = paste0("cd ", hdf5_dir, " && autoconf"))
system(command = paste0("cd ", hdf5_dir, " && aclocal"))
system(command = paste0("cd ", hdf5_dir, " && automake"))
unlink(file.path(hdf5_dir, "autom4te.cache"), recursive = TRUE)
```

## SZIP

Now we carry out a similar process for the SZIP library.

```{r}
szip_source <- tempfile()
download.file(url = "https://support.hdfgroup.org/ftp/lib-external/szip/2.1.1/src/szip-2.1.1.tar.gz",
              dest = szip_source)
untar(tarfile = szip_source, exdir = tempdir())
system2("mv", args = c(file.path(tempdir(), "szip-2.1.1"), file.path(tempdir(), "szip")))
```

```{r}
szip_dir <- file.path(tempdir(), "szip")
```

```{r}
unlink(x = file.path(szip_dir, "test"),
       recursive = TRUE)
```

```{r}
xfun::read_utf8(file.path(szip_dir, "configure.ac")) %>%
  str_remove("test/Makefile") %>%
  xfun::write_utf8(file.path(szip_dir, "configure.ac"))

xfun::read_utf8(file.path(szip_dir, "Makefile.am")) %>%
  str_replace(pattern = "SUBDIRS=src test", replacement = "SUBDIRS=src") %>%
  xfun::write_utf8(file.path(szip_dir, "Makefile.am"))
```

```{r}
system(command = paste0("cd ", szip_dir, " && autoconf"))
system(command = paste0("cd ", szip_dir, " && aclocal"))
system(command = paste0("cd ", szip_dir, " && automake"))
unlink(file.path(szip_dir, "autom4te.cache"), recursive = TRUE)
```

## Create tarball

Move the SZIP source into the HDF5 distribution and create the tarball that will be distributed with the R package.  *We use the system `tar` command here as I can't figure out how to include only the base directory in the tarball when using R's inbuilt function.*

```{r, createTarball}
system2("mv", args = c(szip_dir, file.path(hdf5_dir, "szip")))
system2("tar", args = c("-C", tempdir(), "-czf", file.path(tempdir(), "hdf5small_cxx_1.10.5.tar.gz"), "hdf5"))
```

Finally copy this to the `/tmp/` so it is easy to find, inspect, and move to the package directory if correct.

```{r}
if(file.exists("/tmp/hdf5small_cxx_1.10.5.tar.gz")) { file.remove("/tmp/hdf5small_cxx_1.10.5.tar.gz") }
file.copy(file.path(tempdir(), "hdf5small_cxx_1.10.5.tar.gz"), to = "/tmp/")
```

